\thesistitle

### 2019-08-25 sections/background.tex
Background again again

\section{Internal and external classes}
In this study we differentiate between \textit{internal} and \textit{external} classes. Classes are the components of a software system that contain its functionality. When analyzing a software system, \textit{Internal classes} are classes that belong to this software system specifically, and its source code is included in the project.

\textit{External classes} are classes that a software system uses but do not belong to this specific software system (but to an external dependency). Most often these external dependencies are referenced in a file that its build automation system uses to fetch a projects' dependencies.

Regarding refactoring, a software system can often not change the source code of its dependencies. This can be a burden when an external dependency does not use proper abstractions that are required by a software system. In this study we consider external classes when choosing a refactoring technique, to be sure not to modify them.

\subsection{Maven}
For this study we perform an analysis of external dependencies, to derive more context for internally used concepts. As these external dependencies are most often not included in a software systems' source code, we must first use the projects' build automation system to gather the dependencies. To limit the scope of this process, we decided to focus on only the \textit{Maven}\footnote{Apache Maven is a build automation system mainly used for Java: \url{https://maven.apache.org/}} build automation system.

Maven is a build automation tool, mainly used with the Java programming language. Maven has a simple ecosystem to configure and fetch the binaries and source code of all software projects a given codebase is dependent on. Maven can also be used to run tests for the systems and to package the project as an executable file.

@@

The AST consists of nodes of the following types (examples are for Java):
\begin{itemize}
  \item \textbf{Declarations}: Method-declaration, class-declaration, etc.
  \item \textbf{Statements}: If-statement, switch-statement, expression-statement, etc.
  \item \textbf{Expressions}: Variables, literals, method calls, variable assignments, etc.
  \item \textbf{Types}: Primitives, reference types, void, etc.
  \item \textbf{Clauses}: Catch-clause, else-clause, etc.
\end{itemize}

@@

\subsubsection{Code Smells}
Code smells are patterns in code that should be avoided because they harm system design. Duplicate code is, among many others, one example of a code smell. Having many code smells present in source code can significantly increase the maintenance costs of the source code or even render a software system unmaintainable. This increases technical debt: a debt present in the system the will have to be repaid at a later point of time. Such technical debt often comes at the expense of developing new features, fixing bugs and other aspects surrounding the functional behavior of a program.

@@

\subsubsection{Maintainability Metrics}
Maintainability metrics are features of source code by which an indication of the maintainability of the source code can be measured. Heitlager at al. \cite{heitlager2007practical} propose a maintainability model that categorize the results of metrics into different ``score'' categories. For instance, with duplication, if 0-3\% of the project is duplicated, the project will get the highest score on a scale of 1 to 5 (only integers, no floating-point numbers). Such a maintainability model is intuitive for measuring the quality of a software system, especially large software systems with a lot of source code to base the metrics on. However, this maintainability model \cite{heitlager2007practical} lacks in measuring fine-grained changes: most small changes will not fall into a different category. Because of that, this model is less suitable for measuring the impact of single refactorings.

\subsection{Java}
We run all our experiments on projects written in Java. Because of that, we get some information that is specific to Java systems. Additionally, we perform automated refactorings on Java systems, which requires the use of Java language concepts. These are explained in this section. Many of these concepts also exist in other programming languages, but often not in all.

\subsubsection{Interfaces}
Java uses the concept of interfaces to loosely couple a specification and its implementation. Interfaces are specifications about what a (group of) objects do. Often, methods don't need an entire object, but just use a few properties of it. Because of that, Java recommends programming to an interface rather than an implementation. This means that we should use abstractions of expected functionality rather than complete implementations of functionality. This can make switching implementations at a later point in time easier.

Often, duplication is the result of inappropriate use of abstractions. Two methods might describe operations on types following the same contract, but because of a lack of abstraction, the programmer may choose to clone such methods. This will result in duplicated methods with minor modifications to match a different implementation. Such duplicates can be mitigated by using proper abstractions, of which one option is to program methods against interfaces (contracts of expected functionality) rather than their full implementations.

\subsubsection{Packages}
Packages (sometimes called ``modules'' in other programming languages) are hierarchical groupings of related concepts. For instance, all UI logic in an application might go into the ``ui'' package. The ``ui'' package can have a subpackage named ``keyhandling'' for all keyboard input related operations. These packages contain package-level declarations, like classes and interfaces. Names of class-level declarations must be unique within the package. %However, duplicate declaration names may exist in subpackages and parent packages.

In Java, all declarations used from outside the package a declaration is in, must be \textit{imported}. Importing a declaration from outside the current package goes by the fully qualified identifier (FQI) of the declaration. The FQI of a declaration gives a unique identifier for a symbol in a codebase. For instance, if our ``keyhandling'' package would contain a ``KeyHandler'' class, its fully qualified identifier would be ``ui.keyhandling.KeyHandler''.

\subsubsection{Visibility}
In Java, we can protect declarations from being used in scopes in which they are supposed to be used. If a method should only be used within the class itself, we can give it a ``private'' visibility. If a method is to be used by its children in its inheritance hierarchy, we can set its visibility to ``protected''. If a method is supposed to be part of the public contract of an object, we can set it to ``public'' visibility. Alternatively, there is ``package'' visibility, for declarations that should only be referenced within the same package. This allows control over what parts of an application should be able to access what declarations.

>>>>

\section{Internal and external classes}
In this study we differentiate between \textit{internal} and \textit{external} classes. Classes are the components of a software system that contain its functionality. When analyzing a software system, \textit{Internal classes} are classes that belong to this software system specifically, and their source code is included in the project.

\textit{External classes} are classes that a software system uses but do not belong to this specific software system (but to an external dependency). Most often these external dependencies are referenced in a file that its build automation system uses to fetch these external libraries.

Regarding refactoring, a software system can often not change the source code of its dependencies. This can be a burden when an external dependency does not use proper abstractions that are required by a software system. In this study, we consider external classes when choosing a refactoring technique, to be sure not to modify them.

\subsection{Maven}
In this study, we perform an analysis of external dependencies, to derive more context for internally used concepts. As these external dependencies are most often not included in a software systems' source code, we must first use the projects' build automation system to gather the dependencies. To limit the scope of this process, we decided to focus on only the \textit{Maven}\footnote{Apache Maven is a build automation system mainly used for Java: \url{https://maven.apache.org/}} build automation system.

Maven is a build automation tool, mainly used with the Java programming language. Maven has a simple ecosystem to configure and fetch the binaries and source code of all software projects a given codebase is dependent on. Maven can also be used to run unit tests and to package a project as an executable file.

@@

The AST consists of nodes of the following types \cite{tomassetti2017javaparser}:
\begin{itemize}
  \item \textbf{Declarations}: Method-declaration, class-declaration, etc.
  \item \textbf{Statements}: If-statement, switch-statement, expression-statement, etc.
  \item \textbf{Expressions}: Variables, literals, method calls, variable assignments, etc.
  \item \textbf{Types}: Primitives, reference types, void, etc.
  \item \textbf{Clauses}: Catch-clause, else-clause, etc.
\end{itemize}

@@

\subsubsection{Code Smells}
Code smells are patterns in code that should be avoided because they harm system design. Duplicate code is, among many others, one example of a code smell. Having many code smells present in source code can significantly increase the maintenance costs of the source code or even render a software system unmaintainable \cite{heitlager2007practical}. This increases technical debt: a debt present in the system the will have to be repaid at a later point of time \cite{tom2013exploration}. Such technical debt often comes at the expense of developing new features, fixing bugs and other aspects surrounding the functional behavior of a program \cite{kruchten2012technical}.

@@

\subsubsection{Maintainability Metrics}
Maintainability metrics are features of source code by which an indication of the maintainability of the source code can be measured. Heitlager et al.~\cite{heitlager2007practical} propose a maintainability model that categorize the results of metrics into different risk profiles. For instance, with duplication, if 0-3\% of the project is duplicated, the project will get the highest score on a scale of 1 to 5. Such a maintainability model is intuitive for measuring the quality of a software system, especially large software systems with a lot of source code to base the metrics on. However, this maintainability model~\cite{heitlager2007practical} lacks in measuring fine-grained changes: most small changes will not fall into a different category. Because of that, this model is less suitable for measuring the impact of single refactorings.

\subsection{Java}
We run all our experiments on projects written in Java. Because of that, we get some information that is specific to Java systems. Additionally, we perform automated refactorings on Java systems, which requires the use of Java language concepts. These are explained in this section. Many of these concepts also exist in other programming languages.

\subsubsection{Interfaces}
Java uses the concept of interfaces to loosely couple a specification and its implementation. Interfaces are specifications of what a (group of) objects do. Often, methods don't need an entire object, but just use a few properties of it. Because of that, Java recommends programming to an interface rather than an implementation. This means that we should use abstractions of expected functionality rather than complete implementations of functionality. This can make switching implementations at a later point in time easier.

Often, duplication is the result of inappropriate use of abstractions. Two methods might describe operations on types following the same contract, but because of a lack of abstraction, the programmer may choose to clone such methods. This will result in duplicated methods with minor modifications to fit a different context. Such duplicates can be mitigated by using proper abstractions, of which one option is to program methods against interfaces (contracts of expected functionality) rather than their full implementations.

\subsubsection{Packages}
Packages (called ``modules'' in some other programming languages) are hierarchical groupings of related concepts. For instance, all UI logic in an application might go into the ``ui'' package. The ``ui'' package can have a subpackage named ``keyhandling'' for all keyboard input related operations. These packages contain type declarations, like classes and interfaces. The names of these type declarations must be unique within the package. %However, duplicate declaration names may exist in subpackages and parent packages.

In Java, all declarations used from outside the package a declaration is in, must be \textit{imported}. Importing a declaration from outside the current package goes by the fully qualified identifier (FQI) of the declaration. The FQI of a declaration is the unique identifier of a type declaration in a codebase. For instance, if the ``keyhandling'' package would contain a ``KeyHandler'' class, its fully qualified identifier would be ``ui.keyhandling.KeyHandler''.

\subsubsection{Visibility}
In Java, we can protect declarations from being used in scopes in which they are not supposed to be used. If a method should only be used within the class itself, we can give it a ``private'' visibility. If a method is to be used by its children in its inheritance hierarchy, we can set its visibility to ``protected''. If a method is supposed to be part of the public contract of an object, we can set it to ``public'' visibility. If a method should be accessible to all types in its package, we can give it a ``package'' visibility. This allows control over what parts of an application should be able to access what declarations.
